\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-19 paper/paper.tex
Updates to paper.

The sequence of operations when a packet is received from the client however
needs careful design for efficient software packet switching. To illustrate how
this is achieved, Figure~\ref{fig:packet_layout} shows how Cyclone manipulates
packet layouts across the two steps of prepending a RAFT header and transmitting
to follower replicas. DPDK describes packets using an ``mbuf'' data
structure. Roughly speaking, an ``mbuf'' consists of a flat array of bytes
actually containing the byte and a fixed size piece of external metadata that
describes various aspects of the packet, most crucially a pointer to the start
and end of the packet in the byte array. DPDK's userspace drivers receive
packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We prepend RAFT related
information specific to each log entry in the headroom and shift the start
pointer appropriately - these operations are standard enough for software packet
switches that DPDK provides convenient library calls to do it. For the final
step, we need to prepare the packet for transmission to the various follower
replicas. To do this we prepare a different packet containing an ethernet header
for the targeted replica and ``chain'' the data packet to each of these
headers. Each header is then separately handed off to the driver for
transmission via the NIC, carrying the data packet with it by association.

@@

\subsubsection{Batching}

>>>>

The sequence of operations when a packet is received from the client however
needs careful design for efficient software packet switching. To illustrate how
this is achieved, Figure~\ref{fig:packet_layout} shows how Cyclone manipulates
packet layouts across the two steps of prepending a RAFT header and transmitting
to follower replicas. DPDK describes packets using an ``mbuf'' data
structure. Roughly speaking, an ``mbuf'' consists of a flat array of bytes
actually containing the byte and a fixed size piece of external metadata that
describes various aspects of the packet, most crucially a pointer to the start
and end of the packet in the byte array. DPDK's userspace drivers receive
packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately -
these operations are standard enough for software packet switches that DPDK
provides convenient library calls to do it. For the final step, we need to
prepare the packet for transmission to the various follower replicas. To do this
we prepare a different packet containing an ethernet header for the targeted
replica and ``chain'' the data packet to each of these headers. Each header is
then separately handed off to the driver for transmission via the NIC, carrying
the data packet with it by association.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

\subsubsection{Batching}
Although the control plane code only runs operations related to the RAFT state
machine it is still slow enough relative to the dataplane to add significant
overhead for each packet. We tackle this problem by applying another classic
software packet switching technique: batching - illustrated in
Figure~\ref{fig:batching}. We use a burst receive call available in the DPDK
userspace driver to receive a burst of client packets at a time. We then chain
these packets together and treat them as a single log entry from the perspective
of RAFT, amortizing the control plane overheads over the packets (at most 32 at
a time due to current driver limitations).

We note there that batching in Cyclone does not involve a latency-throughput
tradeoff like in many other systems~\cite{ix-dataplane}. The batch receive call
we use in DPDK returns immediately with whatever number of packets is available,
including zero. We always flush the transmit buffer after every call to DPDK to
transmit packets to replicas. Therefore, we never tradeoff latency for
throughput when batching.
